The zero--one laws of Kolmogorov and Hewitt--Savage in categorical probability

### 2019-10-30 categorical-zero-one.tex
Drafted intro

\todo[inline]{Add some motivation and summary}

@@

@@

\section{Background on Markov categories}

>>>>

\emph{Markov categories} are an approach to probability and statistics based on category theory, developed by the first author in \cite{markov_cats}, building on much previous work by other authors (see \Cref{background}).
The basic idea is to axiomatize the structure and properties of ``categories of stochastic processes'' - for instance, the category $\Stoch$ where objects are measurable spaces and morphisms are Markov kernels (hence the name).
The goal is to state and prove results from probability and statistics in this context.
This represents a \emph{synthetic} approach to probability, in contrast with the usual measure-theoretic approach (which can be called \emph{analytic}).
Instead of defining \emph{what} the words probability, distribution, random variable, etc, \emph{mean}, we instead state by which operations these objects can be combined and related to each other, and which properties they are supposed to satisfy.
We can draw an analogy to the Euclidean approach to geometry - defining certain basic relations between the geometric objects at hand, and setting down axioms about them - with the more modern description of geometric objects as (very roughly) certain subsets of $\bR^n$.

@@

Very roughly, the Kolmogorov zero--one law states that, given an infinite collection of independent random variables, and an event which is in the $\sigma$-algebra generated by the variables, which is independent of any finite subset of variables, the event has probability one or zero.

The Hewitt--Savage zero--one law is similar - here the variables are supposed to be independent and indetically distributed, and the event is supposed to be invariant under finite permutations of the variables. Again, under these assumptions, the event has probability one or zero.

To formulate these statements in the general context of Markov categories, we need a notion of infinite family of random variables.
To this end, we define a notion of \emph{infinite tensor product}. Given such a tensor product, satisfying a certain condition of compatibility with the structure of the Markov category (\Cref{defn_kolmogorov_ext}), we prove abstract versions of the zero--one laws, \Cref{thm:kolmog} and \Cref{thm:hewsav}.

Last, we apply our results to the case of measurable spaces and Markov kernels, deriving the zero--one laws for random variables in standard Borel spaces (this regularity assumption is needed to make the necessary infinite tensor products exist).

@@

\section{Background on Markov categories}\label{background}
